---
title: "Final Project"
author: "Lilach Herzog & Leslie Cohen" 
date: "13 9 2022" 
output: 
  pdf_document: 
    fig_caption: yes
    number_sections: yes
  word_document: 
    toc: yes
    toc_depth: 2
editor_options: 
  chunk_output_Type: inline
---


Our whole project and code can be found on github at <https://github.com/LeslieLebon/assignment-2.git>, and  the dataset was taken from https://www.kaggle.com/datasets/azmatsiddique/glaucoma-dataset 

Our objective in this project was to analyze the "GlaucomaM" dataset in order to be able to look at a person's data and according to that predict with a certain probability whether or not he
might get sick with glaucoma (so the 'Class' of the seed is the column used for labeling and
classifying).

First, we need to load libraries and our dataset.

# preproccessing

## Upload data and libraries

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(gmodels)
library(C50)
library(class) 
library(tidymodels) # for the rsample package, along with the rest of tidymodels
library(ROCR)
library(reshape2)
#library(ggpmisc)
library(ggplot2)
library(gridExtra)
library(pROC)
library(ggfortify)
library(corrplot)
library(tidyverse)    # data manipulation and visualization
library(kernlab)      # SVM methodology
library(e1071)        # SVM methodology
library(ISLR)         # contains example data set "Khan"
library(RColorBrewer) # customized coloring of plot
library(dplyr)
library(knitr)
library(kableExtra)
library(caret)

glaucoma <- read.csv("GlaucomaM.csv")# read data
```

## look at the data and initial preproccessing
We would like to have a look on our dataset.

```{r, message=FALSE, warning=FALSE, results='hide'}
head(glaucoma)
str(glaucoma)
colSums(glaucoma==0)
glaucoma%>%select((everything()))%>%summarise_all(funs(sum(is.na(.)))) # check NA's
```
We have 196 objects and 63 variables (columns). We looked at the data and saw that there are no NA's we need to take care of, and all columns are numerical, so there is no need to work on that. We can also observe that we have few 0s (between 0 and 48(feature vast)), though that is not an absence of information, but in the range of options. We decided to keep all features for the moment.
We took a quick look at the data as a whole, and specifically the 'Class' column. In order to further understand the data and be able to work with it we also used the "table" function, which builds a contingency table of the counts at each combination of factor levels.

```{r, message=FALSE, warning=FALSE, results='hide'}
print(summary(glaucoma$Class))
table(glaucoma$Class) # quick look specifically at the 'Class' column
```

We want to classify our data according to the 'Class' column, which contains the information about whether all the information about a person (represented in a row) belongs to a healthy (normal) or sick (glaucoma) patient, divided neatly in half. Before starting on teaching the algorithm, we converted the 'Class' column to a factor.
```{r, message=FALSE, warning=FALSE, results='hide'}
glaucoma$Class<-as.factor(glaucoma$Class)
```
Since the whole table is already shuffled according to the 'Class' column, we don't need to shuffle the whole table again
In order to get consistent results we initialized a pseudo random number generator. We also made sure to shuffle our data.
```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
set.seed(123)
glaucoma <- glaucoma[sample(nrow(glaucoma)),]
```

# EDA analysis

Results obtained with corrplot: Positive correlations are displayed in a
blue scale while negative correlations are displayed in a red scale. we
can observe 8 clusters of features that highly correlate. Results obtained with heatmap: we
can observe one big cluster that could be divided in 2 and 2 others
clusters of features.

We would like to observe if we can  cluster patients healthy or with glaucoma with all the features.
We performed first a PCA on all fetures and values.

```{r, message=FALSE, warning=FALSE, results='hide', echo=FALSE}

# Preform PCA on our data

gl.pca <- prcomp(glaucoma[,1:62],
                   center = TRUE,
                   scale. = TRUE)
  
# summary of the 
# prcomp object
summary(gl.pca)
gl_pca_plot <- autoplot(gl.pca,
                          data = glaucoma,
                          colour = 'Class')
  

```
```{r, message=FALSE, warning=FALSE, echo=FALSE}
gl_pca_plot
```

In the PCA, we can observe 2 clusters of patients: glaucoma patients and normal
patients. Unfortunately, some of them were not clustered correctly (mixed clouds of blue and red points) in the same area. This mixture could be, probably to the fact that many features correlate, so our next course of action was to check for any correlation between features to determine which feature are correlated by using pearson correlation

```{r, message=FALSE, warning=FALSE, results='hide'}
#calculation of correlation, using pearson correlation
gl.cor = cor(glaucoma[,c(1:62)], method = c("pearson"))

corrplot(gl.cor)

palette = colorRampPalette(c("green", "white", "red")) (20)
heatmap(x = gl.cor, col = palette, symm = TRUE)

```

```{r, message=FALSE, warning=FALSE, results='hide'}
library("Hmisc")
#This generates one table of correlation coefficients (the correlation matrix) and another table of the p-values. By default, the correlations and p-values #are stored in an object of class type rcorr. To extract the values from this object into a useable data structure, you can use the following syntax:
gl.rcorr = rcorr(as.matrix(glaucoma[,c(1:62)]))
gl.rcorr
gl.coeff = gl.rcorr$r

cor_matrix_rm <- gl.cor                  # Modify correlation matrix
cor_matrix_rm[upper.tri(cor_matrix_rm)] <- 0
diag(cor_matrix_rm) <- 0
cor_matrix_rm

data<-glaucoma[,c(1:62)]
data_new <- data[ , !apply(cor_matrix_rm,    # Remove highly correlated variables
                           2,
                           function(x) any(x > 0.8))]
head(data_new)

colSums(glaucoma[,c(1:62)] != 0)

```
```{r, message=FALSE, warning=FALSE, echo=FALSE}
gl.cor2 = cor(data_new, method = c("pearson"))
corrplot(gl.cor2)
heatmap(x = gl.cor2, col = palette, symm = TRUE)


```

```{r, message=FALSE, warning=FALSE, echo=FALSE}

########
# Preform PCA on our new data

gl.pca2 <- prcomp(data_new,
                   center = TRUE,
                   scale. = TRUE)#play with options
  
```
```{r, message=FALSE, warning=FALSE, results='hide'}
# summary of the 
# prcomp object
summary(gl.pca2)
```
```{r, message=FALSE, warning=FALSE, echo=FALSE}
gl_pca_plot2 <- autoplot(gl.pca2,
                          data = glaucoma,
                          colour = 'Class')
gl_pca_plot2
gl_pca_plot
```

## split into training and test sets

### Splitting

We split the whole data set into 80% training set and 20% test set.
```{r, message=FALSE, warning=FALSE}
index <- sample(nrow(glaucoma),round(nrow(glaucoma)*.8),replace = FALSE) #80% of dataset for training
train_set <- glaucoma[index,]
test_set<- glaucoma[-index,]
train_set_labels <-train_set["Class"] #save true "Classes" of the training set
test_set_labels <-test_set["Class"] #save true "Classes" of the test set
```
Since we know that there are about a half of each Class in the data set, we wanted to be sure that the distribution is as we defined (about 1/2 of each Class in both training and test sets should be 'normal' and 1/2 'glaucoma').

```{r, message=FALSE, warning=FALSE, echo=FALSE}
prop.table(table(train_set$Class))
prop.table(table(test_set$Class))
```
The training set is what we will use to train the algorithm, to help it learn our data, and then we will run the algorithm on 20% of the data (test set), and predict according to what it learned in the training set, and since it is still our data we have the true class each individual belongs to, so we can evaluate how well the algorithm succeeded in predicting.

We then tried insteead of splitting it out own to try the 'initial_split' function, with the buikt in 'training' and 'testing' functions
```{r, message=FALSE, warning=FALSE, echo=FALSE}
data <- initial_split(glaucoma, strata = Class )
data_train <- training(data) # we used the default parameters of splitting size
data_test <- testing(data)
prop.table(table(data_train$Class))
prop.table(table(data_test$Class))
```
instead of 0.5159236/ 0.4840764 partition of glaucoma/ normal in the training set and 0.4358974/0.5641026 in the test set, we got exactly 0.5 in both sets.

Since in most algorithms it is important for the data to be scaled, we decided to start by normalizing it.

## normalization

We decided to normalize the data, thinking that the range of values of each parameter is similar in order to compare the distances of different features with different scales.

```{r, results='hide', message=FALSE, warning=FALSE}
normalize <- function(x) {return ((x - min(x)) / (max(x) - min(x)))}
normalize(c(1, 2, 3, 4, 5)) # test our function 1
normalize(c(10, 20, 30, 40, 50)) # test our function 2
train_set_norm <- as.data.frame(lapply(data_train[,-63], normalize))
test_set_norm <- as.data.frame(lapply(data_test[,-63], normalize))
summary(train_set_norm$ag) # test our dataset
```

# KNN algorithm

In the KNN Algorithm, k is a user-defined constant, and an unlabeled vector (a query or test point) is classified by assigning the label which is most frequent among the k training samples nearest to that query point.

## run the KNN algorithm 
### cross validation
We trained the algorithm on the normalized training set. Instead of choosing a random K to run the algorithm with, we used the 'train' function of the 'caret' library with different options of K (in the range of 1-30)
```{r message=FALSE, warning=FALSE}
train_with_knn_cv<-train(train_set_norm,data_train[,63],
                       method = "knn",
                       tuneGrid = data.frame(k=seq(1,30,2)),
                       trControl = trainControl(method = "cv", number = 10),
                       metric = "Kappa")
```
```{r message=FALSE, warning=FALSE, paged.print=TRUE, echo=FALSE}
print(paste0("the best k for cross validation was: ",train_with_knn_cv[["bestTune"]][["k"]])) # to know which k gave the best result
```
We used confusionMatrix to check how well we did (checked on both the train and the test sets).
```{r message=FALSE, warning=FALSE}
knn_predictions_cv<-predict(train_with_knn_cv, test_set_norm, type = "raw")
confMat_cv <- confusionMatrix(knn_predictions_cv,data_test[,63])
```
We then calculated the statistics
```{r message=FALSE, warning=FALSE}
knn_df_norm_cv<-data.frame(predictions=c(as.numeric(knn_predictions_cv)),labels = c(data_test[,63]))
pROC_norm_knn_cv <- roc(knn_df_norm_cv$labels,knn_df_norm_cv$predictions,
            smoothed = TRUE,
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            plot=FALSE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)
sens.ci <- ci.se(pROC_norm_knn_cv)
knn_cv_stats <- data.frame(method="cross validation", 
                           Accuracy=confMat_cv[["overall"]][["Accuracy"]],
                           AreaUndertheCurve=pROC_norm_knn_cv$auc,
                           Sensitivity=confMat_cv[["byClass"]] [["Sensitivity"]], 
                           Precision=confMat_cv[["byClass"]] [["Precision"]], 
                           Specificity=confMat_cv[["byClass"]] [["Specificity"]])
#grid.arrange(tableGrob(confMat_cv[["table"]]), tableGrob(knn_cv_stats),ncol=1,nrow=2)
print(confMat_cv[["table"]])
print.data.frame(knn_cv_stats, right=FALSE)
```
The sensitivity and recall were both 84%, which is not too bad, but the accuracy was 72%, which is not good, and the precision just 67.7%. 
We wanted to see if using repeated cross validation will yield better results
### repeated cross validation
```{r, message=FALSE, warning=FALSE}
train_with_knn_repeatedcv<-train(train_set_norm,data_train[,63],
                       method = "knn",
                       tuneGrid = data.frame(k=seq(1,30,2)),
                       trControl = trainControl(method = "repeatedcv", number = 10, repeats = 10),
                       metric = "Kappa")
```
```{r message=FALSE, warning=FALSE, paged.print=TRUE, echo=FALSE}
print(paste0("the best k for cross validation was: ",train_with_knn_repeatedcv[["bestTune"]][["k"]])) # to know which k gave the best result
```
Again we used confusionMatrix to check how well we did.
```{r, message=FALSE, warning=FALSE, echo=FALSE}
knn_predictions_rep_cv<-predict(train_with_knn_repeatedcv, test_set_norm, type = "raw")
confMat_rep_cv <- confusionMatrix(knn_predictions_rep_cv,data_test[,63])
knn_df_norm_rep_cv<-data.frame(predictions=c(as.numeric(knn_predictions_rep_cv)),labels = c(data_test[,63]))
pROC_norm_knn_rep_cv <- roc(knn_df_norm_rep_cv$labels,knn_df_norm_rep_cv$predictions)
knn_repeatedcv_stats <- data.frame(method="repeated cross validation", 
                           Accuracy=confMat_rep_cv[["overall"]][["Accuracy"]],
                           AUC=pROC_norm_knn_rep_cv$auc,
                           Sensitivity=confMat_rep_cv[["byClass"]] [["Sensitivity"]], 
                           Precision=confMat_rep_cv[["byClass"]] [["Precision"]], 
                           Specificity=confMat_rep_cv[["byClass"]] [["Specificity"]]) 

knn_cv_stats[nrow(knn_cv_stats)+1,]=knn_repeatedcv_stats
#grid.arrange(tableGrob(confMat_rep_cv[["table"]]), tableGrob(knn_cv_stats),ncol=1,nrow=2)
print(confMat_rep_cv[["table"]])
print.data.frame(knn_cv_stats, right=FALSE)
```
We saw no difference. We also checked if z-score (instead of normalization) makes a difference
### z-score standardization 
```{r knn z score, echo=FALSE, message=FALSE, warning=FALSE}
train_set_z <- as.data.frame(scale(data_train[,-63]))
test_set_z <- as.data.frame(scale(data_test[,-63]))

train_with_knn_repeatedcv_z<-train(train_set_z,data_train[,63],
                       method = "knn",
                       tuneGrid = data.frame(k=seq(1,30,2)),
                       trControl = trainControl(method = "repeatedcv", number = 10, repeats = 10),
                       metric = "Kappa")
print(paste0("the best k for repeatedcv with z score is: ",train_with_knn_repeatedcv_z[["bestTune"]][["k"]])) 
train_cconfMat_rep_cv_z <- confusionMatrix(predict(train_with_knn_repeatedcv_z, train_set_z, type = "raw"),data_train[,63])
knn_predictions_rep_cv_z<-predict(train_with_knn_repeatedcv_z, test_set_z, type = "raw")
confMat_rep_cv_z <- confusionMatrix(knn_predictions_rep_cv_z,data_test[,63])

knn_z_table<-CrossTable(x = data_test[,63], y = knn_predictions_rep_cv_z, prop.chisq=FALSE)
df_z<-data.frame(predictions=c(as.numeric(knn_predictions_rep_cv_z)),labels = c(data_test[,63]))
pROC_knn_z <- roc(df_z$labels,df_z$predictions,
            smoothed = TRUE,
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)
knn_repeatedcv_stats_z <- data.frame(method="z score repeated cv", 
                           Accuracy=confMat_rep_cv_z[["overall"]][["Accuracy"]],
                           AUC=pROC_knn_z$auc,
                           Sensitivity=confMat_rep_cv_z[["byClass"]][["Sensitivity"]], 
                           Precision=confMat_rep_cv_z[["byClass"]][["Precision"]], 
                           Specificity=confMat_rep_cv_z[["byClass"]][["Specificity"]]) 

knn_cv_stats[nrow(knn_cv_stats)+1,]=knn_repeatedcv_stats_z
```

## knn conclusions
```{r message=FALSE,warning=FALSE}
print(confMat_rep_cv[["table"]])
print.data.frame(knn_cv_stats, right=FALSE)
```
We did not see any difference whether we used the crossed validation or the repeated cross validation.
We did see a difference when we used the Z score to scale instead of normalization: we gotincreased accuracy/AUC (from 72% to 84%) and precision (from 67.7% to 90.4%!), but at the cost of the sensitivity/recall (which decreased from 84% to 76%).

# SVM
Since it is important when using SVM to work only on numeric data, and when each feature is scaled, we made sure our data was numeric (all columns originally numeric, and the ‘Class’ column we factorizes), randomized (we used the ‘sample’ function beforehand) and normalized. We built the model:
```{r SVM train, message=FALSE, warning=FALSE}
# create and train model
Class_classifier_linear <-  ksvm(x=train_set$Class, data=data_train[,-63],kernel = "vanilladot")
Class_classifier_linear
```

We can see that the model used 49 support vectors, and that the training error was 0.054795. SVM can be a bit of a black box so we don't know much else.
After building the model on our training set (80% of the data) we ran the model on the test set:
```{r message=FALSE, warning=FALSE}
Class_predictions_linear <- predict(Class_classifier_linear, data_test[,-63])
```
We used the table() function to compare the predicted Class to the true Class in the testing:
```{r message=FALSE, warning=FALSE}
table(Class_predictions_linear, data_test$Class)
confMat_svm_linear <- confusionMatrix(Class_predictions_linear,data_test[,63])
```
```{r message=FALSE, warning=FALSE, results='hide'}
#svm_table_linear<-CrossTable(data_test$Class, Class_predictions_linear, dnn = c('actual Type', 'predicted Type'))
#svm_linear_tbl<-svm_table_linear[["t"]]
#grid.arrange(tableGrob(svm_linear_tbl),ncol=1,nrow=1)
```

Now we wanted to see how well our classifier performed according to the different ways of measuring success (precision, recall, accuracy ans AUC):

```{r message=FALSE, warning=FALSE, echo=FALSE}
df_linear<-data.frame(predictions=c(as.numeric(Class_predictions_linear)),labels = c(data_test$Class))
pROC_svm_linear <- roc(df_linear$labels,df_linear$predictions)
#svm_linear_stats <- data.frame(kernel="linear",precision=precision(svm_linear_tbl), recall= round(recall(svm_linear_tbl),3), accuracy=accuracy(svm_linear_tbl)$.estimate, AUC=pROC_svm_linear$au)
svm_linear_stats <- data.frame(method="SVM linear", 
                           Accuracy=confMat_svm_linear[["overall"]][["Accuracy"]],
                           AUC=pROC_svm_linear[["auc"]],
                           Sensitivity=confMat_svm_linear[["byClass"]] [["Sensitivity"]], 
                           Precision=confMat_svm_linear[["byClass"]] [["Precision"]], 
                           Specificity=confMat_svm_linear[["byClass"]] [["Specificity"]])


#grid.arrange(tableGrob(svm_linear_stats),ncol=1,nrow=1)
#agreement_linear <- Class_predictions_linear == test_set$Class # calculate true/false values
#prop.table(table(agreement_linear)) # true/false preformance percentage
print(confMat_svm_linear[["table"]])
svm_stats<-svm_linear_stats
print.data.frame(svm_stats, right=FALSE)
```
The accuracy is  0.846 and the AUC is 0.857. These are pretty good results. We want to make them even better, but what's really important for us is to minimize the false negatives- type II error- since we are working with information about patients, a result telling a person that might be sick that he is healthy can be much more problematic than a result telling a healthy person that he might be sick, and our recall is 0.944.
We still want to see, maybe a different kernel will do a better job?

We tried using the data that we normalized and compared the 2.
```{r message=FALSE, warning=FALSE, echo=FALSE}
Class_predictions_linear_norm <- predict(Class_classifier_linear, test_set_norm)

table(Class_predictions_linear_norm, data_test$Class)
confMat_svm_linear_norm <- confusionMatrix(Class_predictions_linear_norm,data_test[,63])

df_linear_norm<-data.frame(predictions=c(as.numeric(Class_predictions_linear_norm)),labels = c(data_test$Class))
pROC_svm_linear_norm <- roc(df_linear_norm$labels,df_linear_norm$predictions)
#svm_linear_stats <- data.frame(kernel="linear",precision=precision(svm_linear_tbl), recall= round(recall(svm_linear_tbl),3), accuracy=accuracy(svm_linear_tbl)$.estimate, AUC=pROC_svm_linear$au)
svm_linear_norm_stats <- data.frame(method="SVM linear norm", 
                           Accuracy=confMat_svm_linear_norm[["overall"]][["Accuracy"]],
                           AUC=pROC_svm_linear_norm[["auc"]],
                           Sensitivity=confMat_svm_linear_norm[["byClass"]] [["Sensitivity"]], 
                           Precision=confMat_svm_linear_norm[["byClass"]] [["Precision"]], 
                           Specificity=confMat_svm_linear_norm[["byClass"]] [["Specificity"]])


#agreement_linear <- Class_predictions_linear == test_set$Class # calculate true/false values
#prop.table(table(agreement_linear)) # true/false preformance percentage
print(confMat_svm_linear_norm[["table"]])
svm_stats[nrow(svm_stats)+1,]<-svm_linear_norm_stats
grid.arrange(tableGrob(svm_stats),ncol=1,nrow=1)
```
It looks like using the normalized data actually interferes with the classufucations, so we will use the regular data.
### Improving SVM Model

There are many kernel functions for SVM, but a standard Gaussian Radial
basis function (RBF) kernel is a good place to start. We'll use the
ksvm() function here. Note that since this is a large dataset, running
times can start to add up:

```{r message=FALSE, warning=FALSE, echo=FALSE}
# train svm model with RBF kernel function
Class_classifier_rbf <-  ksvm(x=train_set[,63], data=data_train[,-63],kernel = "rbfdot")
Class_predictions_rbf <- predict(Class_classifier_rbf, data_test[,-63])
confMat_svm_rbf <- confusionMatrix(Class_predictions_rbf,data_test[,63])
#svm_table_rbf<-CrossTable(data_test$Class, Class_predictions_rbf,dnn = c('actual Type', 'predicted Type'))
```

And compare it to the previous model:

```{r message=FALSE, warning=FALSE, echo=FALSE}
# calculate performance

df_rbf<-data.frame(predictions=c(as.numeric(Class_predictions_rbf)),labels = c(data_test$Class))
pROC_svm_rbf <- roc(df_rbf$labels,df_rbf$predictions)
svm_rbf_stats <- data.frame(method="SVM RBF", 
                           Accuracy=confMat_svm_rbf[["overall"]][["Accuracy"]],
                           AUC=pROC_svm_rbf[["auc"]],
                           Sensitivity=confMat_svm_rbf[["byClass"]] [["Sensitivity"]], 
                           Precision=confMat_svm_rbf[["byClass"]] [["Precision"]], 
                           Specificity=confMat_svm_rbf[["byClass"]] [["Specificity"]])
print(confMat_svm_rbf[["table"]])

# compare performance
svm_stats[nrow(svm_stats),]=svm_rbf_stats #run over the norm that is no longer needed
print.data.frame(svm_stats, right=FALSE)

roc_test<-roc.test(pROC_svm_linear,pROC_svm_rbf)
roc_test$estimate

svm_linear_stats[nrow(svm_linear_stats)+1,]=svm_rbf_stats
grid.arrange(tableGrob(svm_linear_stats),ncol=1,nrow=1)

g2 <- ggroc(list(vanilladot=pROC_svm_linear, rbf=pROC_svm_rbf))+labs(title = "linear Vs. RBF")


# And change axis labels to FPR/FPR
linear_roc <- ggroc(pROC_svm_linear, legacy.axes = TRUE)+ ggtitle("linear ROC curve") +geom_segment(aes(x = 0, xend = 1, y = 0, yend = 1), color="darkgrey", linetype="dashed")
rbf_roc <- ggroc(pROC_svm_rbf, legacy.axes = TRUE)+ ggtitle("rbf ROC curve") +geom_segment(aes(x = 0, xend = 1, y = 0, yend = 1), color="darkgrey", linetype="dashed")
grid.arrange(linear_roc + xlab("FPR") + ylab("TPR"),g2+theme(legend.position = "top"),rbf_roc + xlab("FPR") + ylab("TPR"), ncol=3) 

```

We can see a slight decrease in accuracy/AUC (from 88% to 86%), sensitivity (from 84% to 80%) and precision (from 91.3% to 90.9%). In both cases the specificity is 92% which is good.
Other kernels might prove even better, or the cost of constraints parameter C could be varied to modify the width of the decision boundary.

# Random forest

We begin by loading the required libraries:

```{r setup}
library(tidymodels) # for the rsample package, along with the rest of tidymodels
library(modeldata)  # for the cells data
library(vip)        # for variable importance plots
```

We have data for 196 patients, with 63 features. The main outcome variable of interest for us here is called class, which is a factor.Each row is a patient, classified as "glaucoma"  or "normal" . If we can predict these labels accurately, the larger data set can be improved by filtering out the cells most likely to be poorly segmented.


An important aspect when performing prediction is the class balance, lets take a look:

```{r, echo=FALSE, results='hide', message=FALSE}
glaucoma %>% 
  count(Class) %>% 
  mutate(prop = n/sum(n))
```

We have 50% 'glaucoma' patients and 50% 'normal' patients. We can make the splits with the same proportions as the original data (50/50).

## Data Splitting

To split data, we are using the strata argument, which conducts a stratified split. This ensures that, despite the imbalance we noticed in our class variable, our training and test data sets will keep roughly the same proportions of sick andnormal patients as in the original data.

```{r, echo=FALSE, results='hide', message=FALSE}
set.seed(123)
patient_split <- initial_split(glaucoma, 
                            strata = Class)

glaucoma_train <- training(patient_split)
glaucoma_test  <- testing(patient_split)

nrow(glaucoma_train)
nrow(glaucoma_train)/nrow(glaucoma)

# training set proportions by class
glaucoma_train %>% 
  count(Class) %>% 
  mutate(prop = n/sum(n))

# test set proportions by class
glaucoma_test %>% 
  count(Class) %>% 
  mutate(prop = n/sum(n))
```

## Random Forest Modeling

The random Forest algorithm uses many decision trees. For classification, which is what
we are focusing on, the final class result is based on the class that was picked from most trees. 

First we define our model with the parsnip package. Random forests have very little mandatory parameters. 
Here we only define the number of trees:

```{r, echo=FALSE, results='hide', message=FALSE}
# create random forest model
rf_mod <- 
  rand_forest(trees = 1000) %>% 
  set_engine("ranger") %>%   # engine - method of estimation the model will use
  set_mode("classification")
```

We want reproducible results, so we set.seed. Taking a look at the model:

```{r, echo=FALSE, results='hide', message=FALSE}
# train model
library(ranger)
set.seed(234)
rf_fit <- 
  rf_mod %>% 
  fit(Class ~ ., data = glaucoma_train)
rf_fit
```

We can see our model had 1000 trees like we specified, and several other metrics, such as the prediction error, 13% here.

### Estimating performance

Next we want to see if we can improve our model. We do this by changing some things about it. 
The changes can either be random (guessing) or more precise, depending on performance metrics. In our example, we will use the area under the Receiver Operating Characteristic (ROC) curve (which demonstrates the trade-off between the sensitivity and and specificity), and overall classification accuracy.

Using the yardstick package, let's calculate ROC and Accuracy. Notice that we are still only working with the glaucoma_train partition of our data:

```{r, echo=FALSE, results='hide', message=FALSE}
rf_training_pred <- 
  predict(rf_fit, glaucoma_train) %>% 
  bind_cols(predict(rf_fit, glaucoma_train, type = "prob")) %>% 
  # Add the true outcome data back in
  bind_cols(glaucoma_train %>% 
              select(Class))

# ROC calculation
rf_training_pred %>%                # training set predictions
  roc_auc(truth = Class, .pred_glaucoma)

# Accuracy calculation
rf_training_pred %>%                # training set predictions
  accuracy(truth = Class, .pred_class)


```

As we can see, these are very good results. Almost too good....Lets see how the model performs on the test data:

```{r, echo=FALSE, results='hide', message=FALSE}
rf_testing_pred <- 
  predict(rf_fit, glaucoma_test) %>% 
  bind_cols(predict(rf_fit, glaucoma_test, type = "prob")) %>% 
  bind_cols(glaucoma_test %>% select(Class))

# ROC calculation
rf_testing_pred %>%                   # test set predictions
  roc_auc(truth = Class, .pred_glaucoma)

# Accuracy calculation
rf_testing_pred %>%                   # test set predictions
  accuracy(truth = Class, .pred_class)
```

These validation results are lower than the ones we got on the training data.
There are several reasons why training set statistics like the ones shown in this section can be unrealistically optimistic:

- Overfitting - Models like random forests, neural networks, and other black-box methods can essentially memorize the training set. Re-predicting that same set should always result in nearly perfect results.

- The training set does not have the capacity to be a good arbiter of performance. It is not an independent piece of information; predicting the training set can only reflect what the model already knows.

## Resampling

Resampling methods, such as cross-validation and the bootstrap, are empirical simulation systems. They create a series of data sets similar to the training/testing split discussed previously; a subset of the data is used for creating the model and a different subset is used to measure performance. Resampling is always used with the training set. 

Here we'll use 10-fold cross-validation. This means that we'll create 10 "mini" datasets, or folds. We call the majority part of the folds (9 out of 10 in this case) the "analysis set" and the minority the "assessment set". We then train a model using the analysis set, and test it on the assessment set, effectively repeating the modeling process 10 times. This is how its done:

```{r, echo=FALSE, results='hide', message=FALSE}
# create the folds
set.seed(345)
folds <- vfold_cv(glaucoma_train, v = 10)
folds

# here we use a workflow, which bundles together the model specifications and actual modeling
# set the random forest workflow 
rf_wf <- 
  workflow() %>%
  add_model(rf_mod) %>%
  add_formula(Class ~ .)

# add folds to workflow and train model
set.seed(456)
rf_fit_rs <- 
  rf_wf %>% 
  fit_resamples(folds)

rf_fit_rs

collect_metrics(rf_fit_rs)
```

We see these results are lower and look more realistic.
Now, looking at the test set for results, we expect and see much more similar results to the above ROC and accuracy:

```{r, echo=FALSE, results='hide', message=FALSE}
# ROC calculation on test predictions
rf_testing_pred %>%             
  roc_auc(truth = Class, .pred_glaucoma)

# Accuracy calculation on test predictions
rf_testing_pred %>%           
  accuracy(truth = Class, .pred_class)
```


### Tuning hyperparameters

Another way to improve the performance of our models is by changing the parameters we provide them with. This is also called Tuning. Random Forests, as mentioned above, are not very sensitive to such parameters, but decision trees are. Lets see:

```{r, echo=FALSE, results='hide', message=FALSE}
# defining the parameters for the decision tree
tune_spec <- 
  decision_tree(
    cost_complexity = tune(), # to control the size of the tree
    tree_depth = tune()
  ) %>% 
  set_engine("rpart") %>% 
  set_mode("classification")

tune_spec
```

Note that in the above code, tune() is still just a placeholder. 
We will fill in values later on.

Next, we will create several smaller datasets, to try different parameters. 
The grid_regular command below will create 5 such datasets for each parameter combination, meaning 25 in total.

```{r, echo=FALSE, results='hide', message=FALSE}
# create 5 'datasets' with different combinations
tree_grid <- grid_regular(cost_complexity(), 
                          tree_depth(),
                          levels = 5)

tree_grid

tree_grid %>% 
  count(tree_depth)

set.seed(234)
glaucoma_folds <- vfold_cv(glaucoma_train) # create the actual cross-validation folds
```
### Model tuning with a grid

```{r, echo=FALSE, results='hide', message=FALSE}
set.seed(345)
# define the workflow
tree_wf <- workflow() %>%
  add_model(tune_spec) %>%
  add_formula(Class ~ .)

# update and train workflow with grids
tree_res <- 
  tree_wf %>% 
  tune_grid(
    resamples = glaucoma_folds,
    grid = tree_grid
    )

tree_res

tree_res %>% 
  collect_metrics()
```
Its easier to see how the models did with a graph:

```{r, echo=FALSE, results='hide', message=FALSE}
# plot preformance
tree_res %>%
  collect_metrics() %>%
  mutate(tree_depth = factor(tree_depth)) %>%
  ggplot(aes(cost_complexity, mean, color = tree_depth)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) +
  scale_x_log10(labels = scales::label_number()) +
  scale_color_viridis_d(option = "plasma", begin = .9, end = 0)
```

We can see that our tree with a depth of 1, is the worst model according to both metrics and across all candidate values of cost_complexity. Our deepest tree, with a depth of 15, did better.  The show_best() function shows us the top 5 candidate models by default:

```{r, echo=FALSE, results='hide', message=FALSE}
# best tree
tree_res %>%
  show_best("roc_auc")
```

And finally, let's finalize the workflow with the best tree. 

```{r, echo=FALSE, results='hide', message=FALSE}
best_tree <- tree_res %>%
  select_best("roc_auc") # Selecting the best tree

best_tree

# Finalize workflow
final_wf <- 
  tree_wf %>% 
  finalize_workflow(best_tree)

final_wf
```
### Exploring results

So we have our best tree parameters. Lets take a better look:

```{r, echo=FALSE, results='hide', message=FALSE}
final_tree <- 
  final_wf %>%
  fit(data = glaucoma_train) 

final_tree
```

Another way to look at how the model uses the different features is using the vip package:

```{r, echo=FALSE, results='hide', message=FALSE}
#final_tree %>% 
 # pull_workflow_fit() %>% 
  #vip()

final_tree %>% 
  extract_fit_parsnip() %>% 
  vip()
```

## Finalizing the model

Finally, let’s return to our test data and estimate the model performance we expect to see with new data. We can use the function last_fit() with our finalized model; this function fits the finalized model on the full training data set and evaluates the finalized model on the testing data.

```{r, echo=FALSE, results='hide', message=FALSE}
# Set workflow
final_fit <- 
  final_wf %>%
  last_fit(patient_split) 

# train
final_fit %>%
  collect_metrics()

# collect metrics
final_fit %>%
  collect_predictions() %>% 
  roc_curve(Class, .pred_glaucoma) %>% 
  autoplot()

args(decision_tree)
```
We can observe in the ROC curve that in our true positive prediction we have very few false positive predictions.


#for future work- feature selection?
```{r}
roc.list <- roc(Class~., data=glaucoma)
per_feature <- ggroc(roc.list, linetype=2)
per_feature+ ggtitle("per feature")


roc.list <- roc(Class~., data=glaucoma)
g.list <- ggroc(roc.list)
g.list + scale_colour_brewer(palette="RdGy")+ ggtitle("title")
```
